Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)\\ Technical Appendix Including Proofs and Details

### 2017-09-07 fully-abstract-comp/technical-report.tex
added a comment.

The following pseudo program demonstrates the need of opaque capabilities. If we assume a system with no opaque capabilities, then the following programs break compiler correctness
\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
p1 ::= if r1 is length 2 then
         call r1 with the following callback in r5:
           {put $\textit{\texttt{diverging}}$ closure in r2;
            return};
         halt
       else diverge

@@

These design goals drove us to the current definition of our source language and some special design choices.
Particularly, contrary to what one might expect in a language with a native stack, we do not use call/return and push/pop instructions.
The reason is that we are working in a language where the length of code blocks can be observable.
Imagine two trusted blocks of code that looks as follows:
\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
p1:       <assert that r1 contains an executable capability to a block of instr_length (push) + instr_length(jmp)>
          <load return capability to p1-ret1 in r2>
          <load value 0 in r3>
          <jump to r1>
p1-ret1:  <assert that 0 is loaded in the next free stack position>
          <repeat the above with value 1 in r3>
          ...
p1-ret2:  <assert that 1 is loaded in the next free stack position>
          <repeat the above with value 2 in r3>
          ...
p1-ret3:  <assert that 2 is loaded in the next free stack position>
          halt

@@

p1 and p2 each accept an (unprotected) closure from the context whose lenght is restricted.
Together with the checks performed by p1 and p2, there is only one possible implementation of the closure that will not fail, namely
\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
closure: push r3
         jmp r2
\end{lstlisting}

However, this is no longer true in the target language.
Assume that $\mathtt{push r3}$ is implemented as a store to $\rstk$ followed by an increment of $\rstk$.
The adversary can now construct a closure that does not use the standard implementation of push, but instead does this:
\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
evil-closure: jnz r3 (\pcreg+2)
              store \rstk r3
              jmp r2
\end{lstlisting}
\dominique{7-9-2017: the above assumes that we have a way of jnz-ing to pc + 2, which may not be true in our actual assembly, but I haven't managed to construct an example that falls entirely within our language.}
If we now assume that the instructions in evil-closure (jnz+store+jmp) are the same length as the original closure(store+increment+jmp), then the adversary can now construct a closure that discriminates $\mathtt{p1}$ and $\mathtt{p2}$ above.
The essential point here is that an assumption in the source language that relies on facts about instruction length (``a closure restricted to the instruction-size of push+ret that modifies the stack and returns can only be implemented in one way'') is no longer valid, because there are now more low-level ways of interacting with the stack.

To remedy this, we decided to modify our source language to use a different way of interacting with the stack.
This alternative approach stays even closer to the target language.
Specifically, rather than using push/pop and call/return instructions, we do not add any instructions, but instead use special token values that play the role of the stack pointer/return pointer.
Specifically, as shown in \cref{sec:domains}, there are new token capabilities $\stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}$, $\retptrd$ and $\retptrc$ which represent a copy of the stack pointer, and the code and data parts of the sealed return closure.
Rather than invoking push or pop, one would store into the $\stkptr{}$ and modify its current address and as we can see in \cref{sec:source-language}, this has the semantic effect that one would expect.
Similarly, rather than invoking return, one would ccall the $\retptrd$ and $\retptrc$ sealed code/data pair.
Finally, rather than invoking call, \dominique{7-9-2017: err, what's the reason we don't follow the same pattern for call?}.

>>>>

The following pseudo program demonstrates the need of opaque capabilities. If we assume a system with no opaque capabilities, then the following programs break compiler correctness
\begin{lstlisting}
[basicstyle=\sourcecolor{}\ttfamily] 
p1 ::= if r1 is length 2 then
         call r1 with the following callback in r5:
           {put $\textit{\texttt{diverging}}$ closure in r2;
            return};
         halt
       else diverge

@@

These design goals drove us to the current definition of our source language and some special design choices.
Particularly, contrary to what one might expect in a language with a native stack, we do not use call/return and push/pop instructions.
The reason is that we are working in a language where the length of code blocks can be observable.
Imagine two trusted blocks of code that looks as follows:
\begin{lstlisting}
[basicstyle=\sourcecolor{}\ttfamily] 
p1:       <assert that r1 contains an executable capability to a block of 
           instr_length (push) + instr_length(jmp)>
          <load return capability to p1-ret1 in r2>
          <load value 0 in r3>
          <jump to r1>
p1-ret1:  <assert that 0 is loaded in the next free stack position>
          <repeat the above with value 1 in r3>
          ...
p1-ret2:  <assert that 1 is loaded in the next free stack position>
          <repeat the above with value 2 in r3>
          ...
p1-ret3:  <assert that 2 is loaded in the next free stack position>
          halt

@@

p1 and p2 each accept an (unprotected) closure from the context whose lenght is restricted.
Together with the checks performed by p1 and p2, there is only one possible implementation of the closure that will not fail, namely
\begin{lstlisting}
[basicstyle=\sourcecolor{}\ttfamily] 
closure: push r3
         jmp r2
\end{lstlisting}

However, this is no longer true in the target language.
Assume that $\mathtt{push r3}$ is implemented as a store to $\rstk$ followed by an increment of $\rstk$.
The adversary can now construct a closure that does not use the standard implementation of push, but instead does this:
\begin{lstlisting}
[basicstyle=\sourcecolor{}\ttfamily] 
evil-closure: jnz r3 (\pcreg+2)
              store \rstk r3
              jmp r2
\end{lstlisting}
\dominique{7-9-2017: the above assumes that we have a way of jnz-ing to pc + 2, which may not be true in our actual assembly, but I haven't managed to construct an example that falls entirely within our language.}
If we now assume that the instructions in evil-closure (jnz+store+jmp) are the same length as the original closure(store+increment+jmp), then the adversary can now construct a closure that discriminates $\mathtt{p1}$ and $\mathtt{p2}$ above.
The essential point here is that an assumption in the source language that relies on facts about instruction length (``a closure restricted to the instruction-size of push+ret that modifies the stack and returns can only be implemented in one way'') is no longer valid, because there are now more low-level ways of interacting with the stack.

To remedy this, we decided to modify our source language to use a different way of interacting with the stack.
This alternative approach stays even closer to the target language.
Specifically, rather than using push/pop and call/return instructions, we do not add any instructions, but instead use special token values that play the role of the stack pointer/return pointer.
Specifically, as shown in \cref{sec:domains}, there are new token capabilities $\stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}$, $\retptrd$ and $\retptrc$ which represent a copy of the stack pointer, and the code and data parts of the sealed return closure.
Rather than invoking push or pop, one would store into the $\stkptr{}$ and modify its current address and as we can see in \cref{sec:source-language}, this has the semantic effect that one would expect.
Similarly, rather than invoking return, one would ccall the $\retptrd$ and $\retptrc$ sealed code/data pair.
Finally, rather than invoking call, \dominique{7-9-2017: err, what's the reason we don't follow the same pattern for call?}.
\lau{07-09-2017: One would use $\scall{}{}$ with a pair of sealed capabilities to securely jump to this pair - that is to have this pair executed in a new stack frame. The return would happen just by using $\sxjmp{}{}$ with the $\retptrd$ and $\retptrc$. However, this would work with the local capabilities as the real implementation of $\retptrd$ would be a sealed version of the entire stack pointer. In other words, upon jumping back no extra work would have to be done. Now, however, the old stack pointer should be spliced (using the new $\ssplice{}{}$ instruction) with the stack pointer we sent away in order to regain the old stack pointer. In other words, we need something to recognise when a return happens.}
